---
title: 'Station Data vs. Daymet: Climate and impact on model output'
author: "Katie Renwick"
date: "12/23/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=F, warning=F, include=F)
```

```{r}
################################################################################
# Compare Reynold's Creek weather data to Daymet
# 1. Compare station data from site 176
################################################################################
rm(list=ls())
library(tidyverse); theme_set(theme_bw(base_size=12)) # sized for print
library(rgdal)
library(raster)

# paths to data and folder for figures
dpath <- "/Users/poulterlab1/Documents/SageParm/data/"
fpath <- "/Users/poulterlab1/Documents/SageParm/figures/"
cpath <- "/Users/poulterlab1/Documents/LPJGUESS/Climate/Daymet/" # climate data

# color scheme
ggcolor <- c("#80cdc1","#018571", "#a6611a","#dfc27d")

############ Get station data and put in propper format
# Get station data for 2014- sent via email by Gerald Flerchinger, ARS
RC176 <- read.csv(paste(dpath, "ReynoldsC/176.wea.30min.2015.csv", sep=""))

# coordinates of mt big sage site nearby:
ll <- c(43.064483, -116.74862)

# Convert 30-min temp and precip to monthly 
head(RC176)
str(RC176)
RC2 <- RC176 %>% mutate(date=as.Date(datetimec)) %>%
  group_by(date) %>% # group by day to calculate days with precip
  summarise(dayppt=sum(ppta), daytemp=mean(tmp3), daysol=mean(sol)) %>%
  mutate(yearmon=format(date, format="%y-%m")) %>%
  group_by(yearmon) %>%
  summarise(temp=mean(daytemp), ppt=sum(dayppt), meansol=mean(daysol),
            wet=sum(dayppt>0))

head(RC2)
RC2$source <- "Station176"
RC3 <- filter(RC2, yearmon<"15-01")

# Pull in daymet3 data, extract values for points
#Loop through the months and sample the climate data
nmonths <- 24
climateD3 <- matrix(nrow=nmonths,ncol=6)

for(m in 1:nmonths){
  print(paste(round((m/nmonths)*100,2),"% finished"))
  #Sample tmean
  tmean.month.m <- raster(paste(cpath, "tmean_1980-2015.nc4", sep=""), 
                          band=(m+408), varname="tmean")    
  climateD3[m,2] <- raster::extract(tmean.month.m, ll)[1]
  ppt.month.m <- raster(paste(cpath, "prcp_1980-2015.nc4", sep=""), 
                          band=(m+408), varname="prcp")    
  climateD3[m,3] <- raster::extract(ppt.month.m, ll)[1]
  wet.month.m <- raster(paste(cpath, "wetdays_1980-2015.nc4", sep=""), 
                          band=(m+408), varname="wetdays")    
  climateD3[m,5] <- raster::extract(wet.month.m, ll)[1]
  srad.month.m <- raster(paste(cpath, "srad24_1980-2015.nc4", sep=""), 
                          band=(m+408), varname="srad")    
  climateD3[m,4] <- raster::extract(srad.month.m, ll)[1]
} 

climateD32 <- data.frame(climateD3) # weird number issues with matrix
names(climateD32) <- c("yearmon","temp","ppt","meansol","wet","source")
climateD32$yearmon <- RC2$yearmon
climateD32$source <- "Daymet3"

all <- rbind(RC2, climateD32)
head(all)
```

## Impact on model output
### First: impact on variables compared to flux data

```{r}
library(zoo)
detach("package:raster", unload=T) # so select works

# Path to model output:
dpath <- "/Users/poulterlab1/Documents/SageParm/testout/"

# Read in flux data from the tower near mt sage site----------------------------
df <- read.csv(paste(dpath, "../data/ReynoldsC/DailyFluxes2015.csv", sep="")) %>%
  mutate(date=as.Date(day,origin=as.yearmon("2015-01-01"))) %>%
  mutate(yearmon=format(date, format="%b %Y")) %>%
  mutate(Source="Tower") %>%
  mutate(yearmon=as.yearmon(yearmon)) %>%
  group_by(yearmon, Source) 

df_GPP <- df %>%
  summarise(GPP=sum(GPPmbsec)*0.001) %>% # conver to kg
  dplyr::select(yearmon, GPP, Source)

# Note: sign seems reversed for NEE, hence mult. by negative .001  
df_NEE <- df %>%
  summarise(NEE=sum(NEEmbsec)*-0.001) %>% # conver to kg
  dplyr::select(yearmon, NEE, Source) 

df_ET <- df %>%
  summarise(ET=sum(ETmbsec)) %>% # units are mm
  dplyr::select(yearmon, ET, Source)

df_R <- df %>%
  summarise(Rsp=sum(Rspmbsec)*0.001) %>% # conver to kg
  dplyr::select(yearmon, Rsp, Source) 

# Read in all monthly model output data in a loop-------------------------------
vars <- c("maet", "mevap", "mgpp","mintercep","mlai","mnee","mnpp","mpet",
          "mra","mrh","mrunoff","mwcont_lower","mwcont_upper")
modout <- NULL
for(var in vars) {
  s1 <- read.table(paste(dpath, "daymet3_",var,".out", sep=""), header=T) %>%
    gather(Month, value,Jan:Dec) %>%
    mutate(Source="Original", variable=var)
  d31 <- read.table(paste(dpath, "daymet3_test_",var,".out", sep=""), header=T) %>%
    gather(Month, value,Jan:Dec) %>%
    mutate(Source="New", variable=var)
  b <- rbind(s1,d31)
  modout <- rbind(modout,b)
}

modout2 <- modout %>%
  mutate(m=match(Month, month.abb), Year=Year+860) %>%
  mutate(yearmon=as.yearmon(paste(Year,m, sep="-"))) %>%
  filter(yearmon>=2014)

# Read in all annual PFT model output data in a loop-------------------------------
vars <- c("anpp","cmass","dens","fpc","lai")
aout <- NULL
for(var in vars) {
  s1 <- read.table(paste(dpath, "daymet3_",var,".out", sep=""), header=T) %>%
    mutate(Source="Original", variable=var)
  d31 <- read.table(paste(dpath, "daymet3_test_",var,".out", sep=""), header=T) %>%
    mutate(Source="New", variable=var)
  b <- rbind(s1,d31)
  aout <- rbind(aout,b)
}
aout2 <- mutate(aout, Year=Year+860)

# Combine GPP with flux data
GPP1 <- modout2 %>% filter(variable=="mgpp") %>%
  select(yearmon, value, Source) %>%
  rename(GPP=value)
GPP <- rbind.data.frame(GPP1, df_GPP)
# Combine NEE with flux data
NEE1 <- modout2 %>% filter(variable=="mnee") %>%
  select(yearmon, value, Source) %>%
  rename(NEE=value)
NEE <- rbind.data.frame(NEE1, df_NEE)
# Combine respiration with flux data
Rsp1 <- modout2 %>% filter(variable=="mra"|variable=="mrh") %>%
  group_by(yearmon,Source) %>%
  summarise(Rsp=sum(value)) %>%
  select(yearmon, Rsp, Source) 
Rsp <- rbind.data.frame(Rsp1, df_R)
# Combine ET with flux data
ET1 <- modout2 %>% filter(variable=="maet") %>%
  select(yearmon, value, Source) %>%
  rename(ET=value)
ET <- rbind.data.frame(ET1, df_ET)


#############################################
# Check: NEE should equal GPP + R:
df2 <- df %>% mutate(NEE2=-GPPmbsec+Rspmbsec) %>%
  select(ETmbsec,GPPmbsec,Rspmbsec,NEEmbsec,NEE2,yearmon)
head(df2) # Perfect!

# Check guess output: NEE should equal GPP + ra + rh
head(modout2)
mo3 <- modout2 %>% spread(variable,value) %>%
  mutate(NEE2=mgpp-mra-mrh) %>%
  select(maet:mgpp,mra:mrh,mnee,NEE2)
head(mo3) # OK, minus some rounding errors
```

```{r, include=T}
ggplot(data=GPP, aes(x=yearmon, y=GPP, color=Source)) +
  geom_line() +
  geom_point() +
  scale_color_manual(values=ggcolor) +
  ylab("Monthly GPP") +
  xlab("Date")

# Plot NEE
ggplot(data=NEE, aes(x=yearmon, y=NEE, color=Source)) +
  geom_line() +
  geom_point() +
  scale_color_manual(values=ggcolor) +
  ylab("Monthly NEE") +
  xlab("Date")

# Plot respiration
ggplot(data=Rsp, aes(x=yearmon, y=Rsp, color=Source)) +
  geom_line() +
  geom_point() +
  scale_color_manual(values=ggcolor) +
  ylab("Monthly Respiration") +
  xlab("Date")
```

### Second: Plot additional monthly variables

```{r, include=T}
mo3 <- modout %>%
  mutate(m=match(Month, month.abb), Year=Year+860) %>%
  mutate(yearmon=as.yearmon(paste(Year,m, sep="-"))) %>%
  filter(yearmon>=2014) 

ggplot(data=filter(mo3, variable=="mnpp"), 
       aes(x=yearmon, y=value, color=Source)) +
  geom_line() + geom_point() +
  scale_color_manual(values=ggcolor) +
  ylab("Monthly NPP") +
  xlab("Date")

ggplot(data=filter(mo3, variable=="mlai"), 
       aes(x=yearmon, y=value, color=Source)) +
  geom_line() + geom_point() +
  scale_color_manual(values=ggcolor) +
  ylab("Monthly LAI") +
  xlab("Date")

ggplot(data=filter(mo3, variable=="mwcont_upper"), 
       aes(x=yearmon, y=value, color=Source)) +
  geom_line() + geom_point() +
  scale_color_manual(values=ggcolor) +
  ylab("Soil Water Content (upper layer)") +
  xlab("Date")

ggplot(data=filter(mo3, variable=="mwcont_lower"), 
       aes(x=yearmon, y=value, color=Source)) +
  geom_line() + geom_point() +
  scale_color_manual(values=ggcolor) +
  ylab("Soil Water Content (lower layer)") +
  xlab("Date")
```

### Third: plot annual variables by PFT
Note that there is only one year of overlap in the driver data: 2014. I've highlighted this year with a dashed line.

```{r, include=T}
ao3 <- filter(aout2, Year>2007) %>%
  gather(PFT, Value, ARTR:C4)
ggplot(data=filter(ao3, variable=="fpc"), 
       aes(x=Year, y=Value, color=PFT, linetype=Source)) +
  geom_line() + geom_point() +
  ylab("Annual FPC") +
  xlab("Year") +
  geom_vline(xintercept=2014, linetype="dashed")

ggplot(data=filter(ao3, variable=="anpp"), 
       aes(x=Year, y=Value, color=PFT, linetype=Source)) +
  geom_line() + geom_point() +
  ylab("Annual NPP") +
  xlab("Year") +
  geom_vline(xintercept=2014, linetype="dashed")

ggplot(data=filter(ao3, variable=="cmass"), 
       aes(x=Year, y=Value, color=PFT, linetype=Source)) +
  geom_line() + geom_point() +
  ylab("Annual Biomass") +
  xlab("Year") +
  geom_vline(xintercept=2014, linetype="dashed")

ggplot(data=filter(ao3, variable=="dens"), 
       aes(x=Year, y=Value, color=PFT, linetype=Source)) +
  geom_line() + geom_point() +
  ylab("Annual Density") +
  xlab("Year") +
  geom_vline(xintercept=2014, linetype="dashed")

ggplot(data=filter(ao3, variable=="lai"), 
       aes(x=Year, y=Value, color=PFT, linetype=Source)) +
  geom_line() + geom_point() +
  ylab("Annual LAI") +
  xlab("Year") +
  geom_vline(xintercept=2014, linetype="dashed")
```

### Fourth: plot annual variables, looking just at total

```{r, include=T}
ggplot(data=filter(ao3, variable=="fpc"), 
       aes(x=Year, y=Total, color=Source, linetype=Source)) +
  geom_line() + geom_point() +
  scale_color_manual(values=ggcolor) +
  ylab("Annual FPC (Total)") +
  xlab("Year") +
  geom_vline(xintercept=2014, linetype="dashed")

ggplot(data=filter(ao3, variable=="anpp"), 
       aes(x=Year, y=Total, color=Source, linetype=Source)) +
  geom_line() + geom_point() +
  scale_color_manual(values=ggcolor) +
  ylab("Annual NPP") +
  xlab("Year") +
  geom_vline(xintercept=2014, linetype="dashed")

ggplot(data=filter(ao3, variable=="cmass"), 
       aes(x=Year, y=Total, color=Source, linetype=Source)) +
  geom_line() + geom_point() +
  scale_color_manual(values=ggcolor) +
  ylab("Annual Biomass") +
  xlab("Year") +
  geom_vline(xintercept=2014, linetype="dashed")

ggplot(data=filter(ao3, variable=="dens"), 
       aes(x=Year, y=Total, color=Source, linetype=Source)) +
  geom_line() + geom_point() +
  scale_color_manual(values=ggcolor) +
  ylab("Annual Density") +
  xlab("Year") +
  geom_vline(xintercept=2014, linetype="dashed")

ggplot(data=filter(ao3, variable=="lai"), 
       aes(x=Year, y=Total, color=Source, linetype=Source)) +
  geom_line() + geom_point() +
  scale_color_manual(values=ggcolor) +
  ylab("Annual LAI (Total)") +
  xlab("Year") +
  geom_vline(xintercept=2014, linetype="dashed")
```

### Just for fun: NPP over modern times.
This shows what a big effect CO2 has throughout the past century. It's much more obvious with the station data because there is less variation in climate, so the trend in increasing NPP is driven entirely by increasing atmospheric CO2.

```{r, include=T}
# NPP over all of time
ggplot(data=filter(aout2, variable=="anpp", Year>1800), 
       aes(x=Year, y=ARTR, color=Source, linetype=Source)) +
  geom_line() + geom_point() +
  scale_color_manual(values=ggcolor) +
  ylab("Annual NPP") +
  xlab("Year") +
  geom_vline(xintercept=2014, linetype="dashed")
```

## Run parameters, which were the same (except end year) for both simulations:
- param "annual_output" (num 1)
- title "LPJ-GUESS (guess2008) cohort mode"
- vegmode "cohort"
- nyear 1156 !1155 to run to 2014
- ifdailynpp 1
- ifdailydecomp 1 !KE 8.19.14 changed to daily 1 from monthly 0
- ifcalcsla 0
- iffire 1 !fire on 0 fire off 1
- npatch 100 
- patcharea 1000
- estinterval 5 
- ifdisturb 1
- distinterval 100
- ifbgestab 1
- ifsme 1
- ifstochestab 1
- ifstochmort 1
- ifcdebt 1
- ifmanaged 0     
- ifsmoothgreffmort 1
- ifdroughtlimitedestab 1 !KE turned off (0)
- ifrainonwetdaysonly 1
- ifspeciesspecificwateruptake 1
- instype "SWRAD"
- soildata "textfile" 
- climdat "DAYMET" 
- startHis 1120 !1120
- outtype "TEXT"
