---
title: "Reynold’s Creek Flux Comparison"
author: "Katie Renwick"
date: "1/11/2017"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=F, warning=F, include=F)
```

```{r}
################################################################################
# Compare Reynold's Creek weather data to Daymet
# 1. Compare station data from site 176
################################################################################
rm(list=ls())
library(tidyverse); theme_set(theme_bw(base_size=20)) # sized for ppt
library(rgdal)
library(raster)

# paths to data and folder for figures
dpath <- "/Users/poulterlab1/Documents/SageParm/data/"
fpath <- "/Users/poulterlab1/Documents/SageParm/figures/"
cpath <- "/Users/poulterlab1/Documents/LPJGUESS/Climate/Daymet/" # climate data

# color scheme
ggcolor <- c("#80cdc1","#018571", "#a6611a","#dfc27d")

############ Get station data and put in propper format
# Get station data for 2014- sent via email by Gerald Flerchinger, ARS
RC176 <- read.csv(paste(dpath, "ReynoldsC/176.wea.30min.csv", sep=""))

# coordinates of mt big sage site nearby:
ll <- c(43.064483, -116.74862)

# Convert 30-min temp and precip to monthly 
head(RC176)
str(RC176)
RC2 <- RC176 %>% mutate(date=as.Date(datetimec)) %>%
  group_by(date) %>% # group by day to calculate days with precip
  summarise(dayppt=sum(ppta), daytemp=mean(tmp3), daysol=mean(sol)) %>%
  mutate(yearmon=format(date, format="%y-%m")) %>%
  group_by(yearmon) %>%
  summarise(temp=mean(daytemp), ppt=sum(dayppt), meansol=mean(daysol),
            wet=sum(dayppt>0))

head(RC2)
RC2$source <- "Station176"
RC3 <- filter(RC2, yearmon<"15-01")

# Pull in daymet3 data, extract values for points
#Loop through the months and sample the climate data
nmonths <- 24
climateD3 <- matrix(nrow=nmonths,ncol=6)

for(m in 1:nmonths){
  print(paste(round((m/nmonths)*100,2),"% finished"))
  #Sample tmean
  tmean.month.m <- raster(paste(cpath, "tmean_1980-2015.nc4", sep=""), 
                          band=(m+408), varname="tmean")    
  climateD3[m,2] <- raster::extract(tmean.month.m, ll)[1]
  ppt.month.m <- raster(paste(cpath, "prcp_1980-2015.nc4", sep=""), 
                          band=(m+408), varname="prcp")    
  climateD3[m,3] <- raster::extract(ppt.month.m, ll)[1]
  wet.month.m <- raster(paste(cpath, "wetdays_1980-2015.nc4", sep=""), 
                          band=(m+408), varname="wetdays")    
  climateD3[m,5] <- raster::extract(wet.month.m, ll)[1]
  srad.month.m <- raster(paste(cpath, "srad24_1980-2015.nc4", sep=""), 
                          band=(m+408), varname="srad")    
  climateD3[m,4] <- raster::extract(srad.month.m, ll)[1]
} 

climateD32 <- data.frame(climateD3) # weird number issues with matrix
names(climateD32) <- c("yearmon","temp","ppt","meansol","wet","source")
climateD32$yearmon <- RC2$yearmon
climateD32$source <- "Daymet3"

all <- rbind(RC2, climateD32)
head(all)
```

# Station data compared to Daymet
## Temperature

```{r, include=T}
# Plot tmean
ggplot(data=all, aes(x=yearmon,y=temp, group=source,color=source)) +
  geom_point(size=1) +
  geom_line(size=1) +
  scale_color_manual(values=ggcolor, name="Data Source") +
  ylab(expression("Temperature ("*~degree*"C)")) +
  xlab("Date (year-month)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

# Station data compared to Daymet
## Precipitation

```{r, include=T}
# Plot ppt
ggplot(data=all, aes(x=yearmon,y=ppt, group=source,color=source)) +
  geom_point(size=1) +
  geom_line(size=1) +
  scale_color_manual(values=ggcolor, name="Data Source") +
  ylab("Precipitation (mm/month)") +
  xlab("Date (year-month)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

# Station data compared to Daymet
## Wet Days

```{r, include=T}
# Plot wetdays
ggplot(data=all, aes(x=yearmon,y=wet, group=source,color=source)) +
  geom_point(size=1) +
  geom_line(size=1) +
  scale_color_manual(values=ggcolor, name="Data Source") +
  ylab("Wetdays (#/month)") +
  xlab("Date (year-month)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

# Station data compared to Daymet
## Solar Radiation

```{r, include=T}

# Plot srad
ggplot(data=all, aes(x=yearmon,y=meansol, group=source,color=source)) +
  geom_point(size=1) +
  geom_line(size=1) +
  scale_color_manual(values=ggcolor, name="Data Source") +
  ylab("Solar Radiation (w/m2)") +
  xlab("Date (year-month)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

# Mean over 2 years for each variable

```{r, include=T}
# Look at 2-year mean:
all2 <- gather(all, Variable, value, temp:wet)

ggplot(data=all2, aes(x=source, y=value, color=source)) +
  geom_boxplot(notch=T, size=1) +
  scale_color_manual(values=ggcolor, name="Data Source") +
  facet_wrap(~Variable, scale="free")
```

```{r}
library(zoo)
detach("package:raster", unload=T) # so select works

# Path to model output:
dpath <- "/Users/poulterlab1/Documents/SageParm/testout/"

# Read in flux data from the tower near mt sage site----------------------------
df <- read.csv(paste(dpath, "../data/ReynoldsC/DailyFluxes2015.csv", sep="")) %>%
  mutate(date=as.Date(day,origin=as.yearmon("2015-01-01"))) %>%
  mutate(yearmon=format(date, format="%b %Y")) %>%
  mutate(Source="Tower") %>%
  mutate(yearmon=as.yearmon(yearmon)) %>%
  group_by(yearmon, Source) 

df_GPP <- df %>%
  summarise(GPP=sum(GPPmbsec)*0.001) %>% # conver to kg
  dplyr::select(yearmon, GPP, Source)

# Note: sign seems reversed for NEE, hence mult. by negative .001  
df_NEE <- df %>%
  summarise(NEE=sum(NEEmbsec)*-0.001) %>% # conver to kg
  dplyr::select(yearmon, NEE, Source) 

df_ET <- df %>%
  summarise(ET=sum(ETmbsec)) %>% # units are mm
  dplyr::select(yearmon, ET, Source)

df_R <- df %>%
  summarise(Rsp=sum(Rspmbsec)*0.001) %>% # conver to kg
  dplyr::select(yearmon, Rsp, Source) 

# Read in MODIS data for the site, filter for this site-------------------------
modis <- read.csv(paste(dpath, "../data/ReynoldsC/MODIS/lai_gpp.csv", sep="")) %>%
  filter(Latitude==43.064483) %>%
  mutate(yearmon=format(Date, format="%b %Y")) %>%
  mutate(Source="MODIS") %>%
  mutate(yearmon=as.yearmon(yearmon)) %>%
  filter(year>=2014, year<=2015) 

# Read in all monthly model output data in a loop-------------------------------
vars <- c("maet", "mevap", "mgpp","mintercep","mlai","mnee","mnpp","mpet",
          "mra","mrh","mrunoff","mwcont_lower","mwcont_upper")
modout <- NULL
for(var in vars) {
  s1 <- read.table(paste(dpath, "station_",var,".out", sep=""), header=T) %>%
    gather(Month, value,Jan:Dec) %>%
    mutate(Source="Station176", variable=var)
  d31 <- read.table(paste(dpath, "daymet3_",var,".out", sep=""), header=T) %>%
    gather(Month, value,Jan:Dec) %>%
    mutate(Source="Daymet3", variable=var)
  b <- rbind(s1,d31)
  modout <- rbind(modout,b)
}

modout2 <- modout %>%
  mutate(m=match(Month, month.abb), Year=Year+860) %>%
  mutate(yearmon=as.yearmon(paste(Year,m, sep="-"))) %>%
  filter(yearmon>=2014)

# Read in all annual PFT model output data in a loop-------------------------------
vars <- c("anpp","cmass","dens","fpc","lai")
aout <- NULL
for(var in vars) {
  s1 <- read.table(paste(dpath, "station_",var,".out", sep=""), header=T) %>%
    mutate(Source="Station176", variable=var)
  #d1 <- read.table(paste(dpath, "daymet_",var,".out", sep=""), header=T) %>%
    #mutate(Source="Daymet", variable=var)
  d31 <- read.table(paste(dpath, "daymet3_",var,".out", sep=""), header=T) %>%
    mutate(Source="Daymet3", variable=var)
  b <- rbind(s1,d31)
  aout <- rbind(aout,b)
}
aout2 <- mutate(aout, Year=Year+860)

# Combine GPP with flux data
GPP1 <- modout2 %>% filter(variable=="mgpp") %>%
  select(yearmon, value, Source) %>%
  rename(GPP=value)
mgpp <- modis %>% select(yearmon, GPP,Source)
GPP <- rbind.data.frame(GPP1, df_GPP,mgpp) %>%
  mutate(Date=as.Date(yearmon)) %>%
  mutate(Source=factor(Source, levels=c("Daymet3", "Station176","Tower","MODIS"), ordered=TRUE))
# Combine NEE with flux data
NEE1 <- modout2 %>% filter(variable=="mnee") %>%
  select(yearmon, value, Source) %>%
  rename(NEE=value)
NEE <- rbind.data.frame(NEE1, df_NEE) %>%
  mutate(Date=as.Date(yearmon))
# Combine respiration with flux data
Rsp1 <- modout2 %>% filter(variable=="mra"|variable=="mrh") %>%
  group_by(yearmon,Source) %>%
  summarise(Rsp=sum(value)) %>%
  select(yearmon, Rsp, Source) 
Rsp <- rbind.data.frame(Rsp1, df_R)
Rsp$Date <- as.Date(Rsp$yearmon)
# Combine ET with flux data
ET1 <- modout2 %>% filter(variable=="maet") %>%
  select(yearmon, value, Source) %>%
  rename(ET=value)
ET <- rbind.data.frame(ET1, df_ET) %>%
  mutate(Date=as.Date(yearmon))
# Combine LAI with modis data
LAI1 <- modout2 %>% filter(variable=="mlai") %>%
  select(yearmon, value, Source) %>%
  rename(LAI=value)
mlai <- modis %>% select(yearmon, LAI,Source)
LAI <- rbind.data.frame(LAI1,mlai) %>%
  mutate(Date=as.Date(yearmon)) %>%
  mutate(Source=factor(Source, levels=c("Daymet3", "Station176","Tower","MODIS"), ordered=TRUE))

#############################################
# Check: NEE should equal GPP + R:
df2 <- df %>% mutate(NEE2=-GPPmbsec+Rspmbsec) %>%
  select(ETmbsec,GPPmbsec,Rspmbsec,NEEmbsec,NEE2,yearmon)
head(df2) # Perfect!

# Check guess output: NEE should equal GPP + ra + rh
head(modout2)
mo3 <- modout2 %>% spread(variable,value) %>%
  mutate(NEE2=mgpp-mra-mrh) %>%
  select(maet:mgpp,mra:mrh,mnee,NEE2)
head(mo3) # OK, minus some rounding errors
```

# Impact of climate drivers on model output
## GPP

```{r, include=T}
ggplot(data=GPP, aes(x=Date, y=GPP, color=Source)) +
  geom_line(size=1) +
  geom_point(size=1.5) +
  scale_color_manual(values=ggcolor) +
  ylab("Monthly GPP") +
  xlab("Date")
```

# Impact of climate drivers on model output
## NEE

```{r, include=T}
# Plot NEE
ggplot(data=NEE, aes(x=Date, y=NEE, color=Source)) +
  geom_line(size=1) +
  geom_point(size=1.5) +
  scale_color_manual(values=ggcolor) +
  ylab("Monthly NEE") +
  xlab("Date")
```

# Impact of climate drivers on model output
## Respiration

```{r, include=T}
# Plot respiration
ggplot(data=Rsp, aes(x=Date, y=Rsp, color=Source)) +
  geom_line(size=1) +
  geom_point(size=1.5) +
  scale_color_manual(values=ggcolor) +
  ylab("Monthly Respiration") +
  xlab("Date")
```

# Impact of climate drivers on model output
## NPP

```{r, include=T}
mo3 <- modout %>%
  mutate(m=match(Month, month.abb), Year=Year+860) %>%
  mutate(yearmon=as.yearmon(paste(Year,m, sep="-"))) %>%
  filter(yearmon>=2014) %>%
  mutate(Date=as.Date(yearmon))

ggplot(data=filter(mo3, variable=="mnpp"), 
       aes(x=Date, y=value, color=Source)) +
  geom_line(size=1) + geom_point(size=1.5) +
  scale_color_manual(values=ggcolor) +
  ylab("Monthly NPP") +
  xlab("Date")
```

# Impact of climate drivers on model output
## LAI

```{r, include=T}
ggplot(data=LAI, aes(x=Date, y=LAI, color=Source)) +
  geom_line(size=1) +
  geom_point(size=1.5) +
  scale_color_manual(values=ggcolor, drop=F) +
  ylab("Monthly LAI") +
  xlab("Date")
```

```{r, include=T}
# ggplot(data=filter(mo3, variable=="mlai"),
#        aes(x=yearmon, y=value, color=Source)) +
#   geom_line() + geom_point() +
#   scale_color_manual(values=ggcolor) +
#   ylab("Monthly LAI") +
#   xlab("Date")

# ggplot(data=filter(mo3, variable=="mwcont_upper"), 
#        aes(x=yearmon, y=value, color=Source)) +
#   geom_line() + geom_point() +
#   scale_color_manual(values=ggcolor) +
#   ylab("Soil Water Content (upper layer)") +
#   xlab("Date")
# 
# ggplot(data=filter(mo3, variable=="mwcont_lower"), 
#        aes(x=yearmon, y=value, color=Source)) +
#   geom_line() + geom_point() +
#   scale_color_manual(values=ggcolor) +
#   ylab("Soil Water Content (lower layer)") +
#   xlab("Date")
```

# Annual variables by Plant Functional Type (PFT)
## Foliar Projective Cover

```{r, include=T}
ao3 <- filter(aout2, Year>2007) %>%
  gather(PFT, Value, ARTR:C4)
ggplot(data=filter(ao3, variable=="fpc"), 
       aes(x=Year, y=Value, color=PFT, linetype=Source)) +
  geom_line(size=1) + geom_point(size=2) +
  #scale_color_manual(values=ggcolor) +
  ylab("Annual FPC") +
  xlab("Year") +
  geom_vline(xintercept=2014, linetype="dashed")
```

# Net Primary Productivity

```{r, include=T}
ggplot(data=filter(ao3, variable=="anpp"), 
       aes(x=Year, y=Value, color=PFT, linetype=Source)) +
  geom_line(size=1) + geom_point(size=2) +
  #scale_color_manual(values=ggcolor) +
  ylab("Annual NPP") +
  xlab("Year") +
  geom_vline(xintercept=2014, linetype="dashed")
```

# Biomass

```{r, include=T}
ggplot(data=filter(ao3, variable=="cmass"), 
       aes(x=Year, y=Value, color=PFT, linetype=Source)) +
  geom_line(size=1) + geom_point(size=2) +
  #scale_color_manual(values=ggcolor) +
  ylab("Annual Biomass") +
  xlab("Year") +
  geom_vline(xintercept=2014, linetype="dashed")

# ggplot(data=filter(ao3, variable=="dens"), 
#        aes(x=Year, y=Value, color=PFT, shape=Source)) +
#   geom_line() + geom_point() +
#   scale_color_manual(values=ggcolor) +
#   ylab("Annual Density") +
#   xlab("Year") +
#   geom_vline(xintercept=2014, linetype="dashed")
```


```{r, include=T}
# ggplot(data=filter(ao3, variable=="fpc"), 
#        aes(x=Year, y=Total, color=Source, shape=Source)) +
#   geom_line() + geom_point() +
#   scale_color_manual(values=ggcolor) +
#   ylab("Annual FPC (Total)") +
#   xlab("Year") +
#   geom_vline(xintercept=2014, linetype="dashed")

# ggplot(data=filter(ao3, variable=="lai"), 
#        aes(x=Year, y=Total, color=Source)) +
#   geom_line(size=1) + geom_point(size=2) +
#   scale_color_manual(values=ggcolor) +
#   ylab("LAI") +
#   xlab("Year") +
#   geom_vline(xintercept=2014, linetype="dashed")

# ggplot(data=filter(ao3, variable=="cmass"), 
#        aes(x=Year, y=Total, color=Source, shape=Source)) +
#   geom_line(size=1) + geom_point(size=1.5) +
#   scale_color_manual(values=ggcolor) +
#   ylab("Annual Biomass") +
#   xlab("Year") +
#   geom_vline(xintercept=2014, linetype="dashed")

# ggplot(data=filter(ao3, variable=="dens"), 
#        aes(x=Year, y=Total, color=Source, shape=Source)) +
#   geom_line() + geom_point() +
#   scale_color_manual(values=ggcolor) +
#   ylab("Annual Density") +
#   xlab("Year") +
#   geom_vline(xintercept=2014, linetype="dashed")
```

# Next Steps
1. Plant community data- is the problem relative % of sage?
2. Additional sites/years for comparison. Can LPJ-GUESS replicate spatial and temporal trends in vegetation response to climate?
3. Parameter optimization: match flux data to determine new estimates for parameters with limited data (ex. root turnover rate).
4. Learn something about the ecology/physiology and climate sensitivity of sagebrush!