---
title: "Flux Comparison"
author: "Katie Renwick"
date: "1/20/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=F, warning=F, include=F)
```

```{r}
rm(list=ls())
library(tidyverse); theme_set(theme_bw(base_size=12)) # sized for ppt
library(rgdal)
library(raster)
library(zoo) #fucking hate time series!

# paths to data and folder for figures
dpath <- "/Users/poulterlab1/Documents/SageParm/data/"
fpath <- "/Users/poulterlab1/Documents/SageParm/figures/"
cpath <- "/Users/poulterlab1/Documents/LPJGUESS/Climate/Daymet/" # climate data

# color scheme
ggcolor <- c("#80cdc1","#018571", "#a6611a","#dfc27d")

############ Get station data and put in propper format
# Get station data for 2014- sent via email by Gerald Flerchinger, ARS
RC176 <- read.csv(paste(dpath, "ReynoldsC/176.wea.30min.2015.csv", sep="")) %>%
  mutate(Source="Station", Site="mbsec")
RC098c <- read.csv(paste(dpath, "ReynoldsC/098c.wea.30min.2015.csv", sep="")) %>%
  mutate(Source="Station", Site="wbsec")
RC127 <- read.csv(paste(dpath, "ReynoldsC/127.wea.30min.2015.csv", sep="")) %>%
  mutate(Source="Station", Site="losec")
RC138j10 <- read.csv(paste(dpath, "ReynoldsC/138j10.wea.30min.2015.csv", sep="")) %>%
  mutate(Source="Station", Site="h08ec")
RCweather <- rbind(RC176,RC098c,RC127,RC138j10)

# coordinates of flux sites:
sites <- read.csv(paste(dpath, "ReynoldsC/EC-CoreSites-coords.csv", sep=""))
ll <- dplyr::select(sites, Lon,Lat)
Site <- dplyr::select(sites, Site)

# Convert 30-min temp and precip to monthly 
RC2 <- RCweather %>% mutate(date=as.Date(datetimec)) %>%
  group_by(date,Source,Site) %>% # group by day to calculate days with precip
  summarise(dayppt=sum(ppta), daytemp=mean(tmp3), daysol=mean(sol)) %>%
  mutate(yearmon=format(date, format="%y-%m")) %>%
  group_by(yearmon,Source,Site) %>%
  summarise(temp=mean(daytemp), ppt=sum(dayppt), meansol=mean(daysol),
            wet=sum(dayppt>0))

head(RC2)

# Pull in daymet3 data, extract values for points
#Loop through the months and sample the climate data
nmonths <- 24
ClimateD3 <- NULL
TS <- unique(RC2$yearmon)

for(m in 1:nmonths){
  print(paste(round((m/nmonths)*100,2),"% finished"))
  #Sample tmean
  tmean.month.m <- raster(paste(cpath, "tmean_1980-2015.nc4", sep=""), 
                          band=(m+408), varname="tmean")    
  temp <- raster::extract(tmean.month.m, ll)
  ppt.month.m <- raster(paste(cpath, "prcp_1980-2015.nc4", sep=""), 
                        band=(m+408), varname="prcp")    
  ppt <- raster::extract(ppt.month.m, ll)
  wet.month.m <- raster(paste(cpath, "wetdays_1980-2015.nc4", sep=""), 
                        band=(m+408), varname="wetdays")    
  wet <- raster::extract(wet.month.m, ll)
  srad.month.m <- raster(paste(cpath, "srad24_1980-2015.nc4", sep=""), 
                         band=(m+408), varname="srad")    
  meansol <- raster::extract(srad.month.m, ll)
  
  # Get dates
  yearmon <- rep(TS[m], 4)
  all <- cbind(yearmon,Site,temp,ppt,meansol,wet)
  ClimateD3 <- rbind(ClimateD3,all)
} 
ClimateD3 <- as_tibble(ClimateD3)
ClimateD32 <- ClimateD3 %>%
  mutate(Source="Daymet") %>%
  dplyr::select(yearmon, Source, Site, temp:wet) %>%
  mutate(yearmon=as.character(yearmon))
head(ClimateD32)
head(RC2)
all <- rbind.data.frame(RC2, ClimateD32)
all$Source <- as.factor(all$Source)
all$Site <- as.factor(all$Site)
all$Date <- as.Date(as.yearmon(all$yearmon,format="%y-%m"))
head(all)
```

## Compare Daymet vs. Station Data
Note: Several sites are near the corner of a Daymet grid cell. The weather stations are also not exactly at coordinates of the sites. This could contribute to some of the differences in monthly weather.

```{r, include=T}
################################################################################
# Plot sites over Daymet temperature
#convert the raster to points for plotting
map.p <- rasterToPoints(tmean.month.m)
df <- data.frame(map.p) %>%
  filter(x<=-116.7,x>=-116.77, y>=43.05,y<=43.18)
colnames(df) <- c("Longitude", "Latitude", "MAT")

# Plot Daymet grid with study sites shown
ggplot(data=df, aes(y=Latitude, x=Longitude)) +
  geom_raster(aes(fill=MAT)) +
  geom_point(data=sites, aes(x=Lon, y=Lat), color="white", size=2, shape=1) +
  geom_text(data=sites,aes(x=Lon, y=Lat,label=Site),nudge_y = .005, color="white") +
  theme_bw() +
  coord_equal() +
  scale_fill_gradient("Dec. 2015 Temp.") +
  theme(axis.title.x = element_text(size=16),
        axis.title.y = element_text(size=16, angle=90),
        axis.text.x = element_text(size=14),
        axis.text.y = element_text(size=14),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "right",
        legend.key = element_blank())

# Plot tmean
ggplot(data=all, aes(x=Date,y=temp, linetype=Source, color=Site)) +
  geom_point(size=1) +
  geom_line(size=1) +
  ylab(expression("Temperature ("*~degree*"C)")) +
  xlab("Date (year-month)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~Site)

## Precipitation
ggplot(data=all, aes(x=Date,y=ppt, linetype=Source, color=Site)) +
  geom_point(size=1) +
  geom_line(size=1) +
  ylab("Precipitation (mm/month)") +
  xlab("Date (year-month)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~Site)

## Wet Days
ggplot(data=all, aes(x=Date,y=ppt, linetype=Source, color=Site)) +
  geom_point(size=1) +
  geom_line(size=1) +
  ylab("Wetdays (#/month)") +
  xlab("Date (year-month)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~Site)

## Solar Radiation
ggplot(data=all, aes(x=Date,y=ppt, linetype=Source, color=Site)) +
  geom_point(size=1) +
  geom_line(size=1) +
  #scale_color_manual(values=ggcolor, name="Site") +
  ylab("Solar Radiation (w/m2)") +
  xlab("Date (year-month)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~Site)

# Mean over 2 years for each variable
all2 <- gather(all, Variable, value, temp:wet)
ggplot(data=all2, aes(x=Site, y=value, color=Source)) +
  geom_boxplot(notch=T, size=1) +
  scale_color_manual(values=ggcolor, name="Data Source") +
  ylab("Value") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~Variable, scale="free")
```

```{r}
#############################################################################
# Plot climate: station vs. Daymet.
# Plot model output, compared to flux, for 2015
#############################################################################
rm(list=ls())
library(tidyverse); theme_set(theme_bw(base_size=12)) # sized for ppt
library(zoo) #time series
detach("package:raster", unload=T) # so select works

# paths to data and folder for figures
fpath <- "/Users/poulterlab1/Documents/SageParm/figures/"
cpath <- "/Users/poulterlab1/Documents/LPJGUESS/Climate/Daymet/" # climate data

# Path to model output:
dpath <- "/Users/poulterlab1/Documents/SageParm/RCout/"

################################################################################
# 1. READ IN DATA

# Pull in site list to get coordinates
sites <- read.csv(paste(dpath, "../data/ReynoldsC/EC-CoreSites-coords.csv", sep="")) %>%
  mutate(Lon=round(Lon,4), Lat=round(Lat,4))

# Read in flux data from 4 RC sites
sitelist <- c("wbsec","losec","h08ec","mbsec")
df <- read.csv(paste(dpath, "../data/ReynoldsC/DailyFluxes2015.csv", sep="")) %>%
  mutate(date=as.Date(day,origin=as.yearmon("2015-01-01"))) %>%
  mutate(yearmon=format(date, format="%b %Y")) %>%
  mutate(Source="Tower") %>%
  mutate(yearmon=as.yearmon(yearmon)) %>%
  group_by(yearmon, Source) %>%
  select(-jday, -year, -day, -date) %>%
  summarise_each(funs(sum)) %>%
  gather(Site, Value, ETwbsec:Rspmbsec) %>%
  separate(Site, into=c("Variable","Site"), sep=-6) %>%
  mutate(Value=ifelse(Variable=="GPP", Value*0.001, Value)) %>% # conver to kg
  mutate(Value=ifelse(Variable=="NEE", Value*-0.001, Value)) %>% # conver to kg
  mutate(Value=ifelse(Variable=="Rsp", Value*0.001, Value)) # conver to kg
head(df)
unique(df$Site)
unique(df$Variable)

# Read in model output for the 4 sites
# Read in all monthly model output data in a loop-------------------------------
vars <- c("maet", "mevap", "mgpp","mintercep","mlai","mnee","mnpp","mpet",
          "mra","mrh","mrunoff","mwcont_lower","mwcont_upper")
modout <- NULL
for(var in vars) {
  s1 <- read.table(paste(dpath, "daymet3_",var,".out", sep=""), header=T) %>%
    gather(Month, value,Jan:Dec) %>%
    mutate(Source="Model", variable=var)
  modout <- rbind(modout,s1)
}

modout2 <- modout %>%
  mutate(m=match(Month, month.abb), Year=Year+860) %>%
  mutate(yearmon=as.yearmon(paste(Year,m, sep="-"))) %>%
  filter(yearmon>=2014)
head(modout2)
table(modout2$Lon, modout2$Lat)
table(sites$Lon, sites$Lat)
modout3 <- merge(modout2, sites, by=c("Lon","Lat")) %>%
  select(yearmon, Source,variable,Site,value) %>%
  rename(Value=value, Variable=variable) %>%
  mutate(Variable=ifelse(Variable=="mgpp","GPP", Variable)) %>%
  mutate(Variable=ifelse(Variable=="mnee","NEE", Variable)) %>%
  spread(Variable, Value) %>%
  mutate(Rsp=mra+mrh, ET=mevap+maet, NEE=NEE*-1) %>%
  gather(Variable,Value,GPP:ET)
head(modout3)

# Read in all annual PFT model output data in a loop----------------------------
vars <- c("anpp","cmass","dens","fpc","lai")
aout <- NULL
for(var in vars) {
  s1 <- read.table(paste(dpath, "daymet3_",var,".out", sep=""), header=T) %>%
    mutate(Source="Model", variable=var)
  aout <- rbind(aout,s1)
}
aout2 <- mutate(aout, Year=Year+860)

# Combine flux data with model output-------------------------------------
df2 <- data.frame(df)
flux <- rbind.data.frame(df2,modout3) %>%
  filter(yearmon>=min(df2$yearmon))
flux$Date <- as.Date(as.yearmon(flux$yearmon,format="%y-%m"))
head(flux)
```

## Compare Modeled (Daymet drivers)  vs. Tower Flux Data
Note that for GPP, the model seems to under-estimate the peak and over-estimate the fall values. I think this may be partly due to the fact that sagebrush drops it's ephemeral leaves ~June, so SLA changes and leaf cover is reduced (in nature). In the model sagebrush is evergreen.

```{r, include=T}
################################################################################
# Make Plots
ggplot(data=filter(flux, Variable=="GPP"), 
       aes(x=Date, y=Value, linetype=Source, color=Site)) +
  geom_line() +
  geom_point() +
  #scale_color_manual(values=ggcolor) +
  ylab("Monthly GPP") +
  xlab("Date") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~Site)

# Plot NEE
ggplot(data=filter(flux, Variable=="NEE"), 
       aes(x=Date, y=Value, linetype=Source, color=Site)) +
  geom_line() +
  geom_point() +
  #scale_color_manual(values=ggcolor) +
  ylab("Monthly NEE") +
  xlab("Date") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~Site)

# Plot respiration
ggplot(data=filter(flux, Variable=="Rsp"), 
       aes(x=Date, y=Value, linetype=Source, color=Site)) +
  geom_line() +
  geom_point() +
  #scale_color_manual(values=ggcolor) +
  ylab("Monthly Respiration") +
  xlab("Date") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~Site)

# Plot ET
ggplot(data=filter(flux, Variable=="ET"), 
       aes(x=Date, y=Value, linetype=Source, color=Site)) +
  geom_line() +
  geom_point() +
  #scale_color_manual(values=ggcolor) +
  ylab("Monthly ET") +
  xlab("Date") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~Site)
```